---
title: "MLR Model"
author: "Preston Dunton, Trevor Overton, Jasmine DeMeyer"
date: "4/24/2022"
output: pdf_document
---

# Imports and Constants

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
source('./gibbs_util.R')

RANDOM_SEED = 440
```

# Data Loading / Cleaning

```{r setup}
load('./burritodata.Rda')
head(burrito)

# Count Remove NA Cost rows
which(is.na(burrito$Cost))
burrito = burrito[!is.na(burrito$Cost),]
nrow(burrito)


burrito<-burrito%>%mutate(Vegetable=as.logical(Pineapple+Bell_peper+Tomato+
                                                 Cabbage+Mushroom+Corn+
                                                 Carrots+Zucchini))
burrito<-burrito%>%mutate(Breakfast=as.logical(Egg+Bacon+Ham))
burrito<-burrito%>%mutate(Other=as.logical(Fish+Taquito+Chile_relleno+
                                             Nopales+Sushi+Lobster))
burrito<-select(burrito, -c('Pineapple', 'Bell_peper', 'Tomato', 'Cabbage', 
                            'Mushroom', 'Corn', 'Carrots', 'Zucchini', 
                            'Egg', 'Ham', 'Fish', 'Taquito', 'Chile_relleno', 
                            'Nopales', 'Sushi', 'Lobster', 'Bacon'))
burrito<-burrito%>%mutate_at(c('Vegetable','Other','Breakfast'),as.double)
head(burrito)

burrito = burrito %>% mutate(Num_Proteins= Chicken + Beef + Pork + Shrimp + Other + Breakfast)

ingredient_cols = colnames(burrito)[18:36]
ingredient_cols

ingredient_X = as.matrix(burrito[ingredient_cols])
dim(ingredient_X); head(ingredient_X)

cost_y = burrito$Cost
length(cost_y); head(cost_y)

```

```{r}
num_burrito_ingredients = c()
for (ingredient in ingredient_cols) {
  num_burrito_ingredients = c(num_burrito_ingredients, 
                              sum(burrito[ingredient]))  
}
ingredient_counts_df = data.frame(ingredient=ingredient_cols, 
                                  count=num_burrito_ingredients)

# sort by count
ingredient_counts_df = ingredient_counts_df[order(ingredient_counts_df$count, decreasing=TRUE),] 

kable(ingredient_counts_df, row.names=FALSE)

barplot(ingredient_counts_df$count, ylab='Number of Burritos',
        main='Ingredient Distribution', 
        names.arg=ingredient_counts_df$ingredient, las=2)
```

# Definitions of Priors and Constants

```{r priors_and_consts}

p = ncol(ingredient_X) + 1
tau_2 = 4
prior_sigma = 1.5
a = 1 / (prior_sigma^4)
b = 1 / (prior_sigma^2)

```

# Model Fit with Gibbs Sampler

```{r gibbs}
set.seed(RANDOM_SEED)

mlr_post_dist = mlr_gibbs(ingredient_X, cost_y, mu=rep(0, p), tau_2, a, b)
mlr_post_dist = mlr_post_dist[5001:1000, ]
summarize_dist(mlr_post_dist, colnames(mlr_post_dist), round_places=2)

```

This model won't work for us because it fits prices to be negative.  Instead, why don't we use a truncated Gibbs sampler.

# Full Truncated Gibbs Model
```{r t_gibbs}
set.seed(RANDOM_SEED)

truncated_post_dist<-truncated_gibbs(ingredient_X, cost_y, mu=rep(1, p), 
                                     tau_2, a, b, lb=rep(0,p), ub=rep(Inf,p))
truncated_post_dist = truncated_post_dist[5001:1000, ]
head(truncated_post_dist)
```

## Model Diagnostics

```{r t_gibbs_trace_plots}
plot_traces(truncated_post_dist[,1:5], 'Parameter Traces (After Burn In)')
plot_traces(truncated_post_dist[,6:10], 'Parameter Traces (After Burn In)')
plot_traces(truncated_post_dist[,11:15], 'Parameter Traces (After Burn In)')
plot_traces(truncated_post_dist[,16:21], 'Parameter Traces (After Burn In)')
```

```{r t_gibbs_acf_plots}
acf_plots(truncated_post_dist[,1:5])
acf_plots(truncated_post_dist[,6:10])
acf_plots(truncated_post_dist[,11:15])
acf_plots(truncated_post_dist[,16:21])
```


```{r t_gibbs_summary}
summarize_dist(truncated_post_dist, colnames(truncated_post_dist), round_places=2)
```

# Protein Model

```{r protein subsetting}

num_burritos_no_protein = sum(burrito$Num_Proteins == 0)
num_burritos_no_protein
num_burritos_double_protein = sum(burrito$Num_Proteins == 2)
num_burritos_double_protein

burrito_no_double_protein = burrito[burrito$Num_Proteins != 2, ]
head(burrito_no_double_protein)

burrito_no_double_protein = burrito_no_double_protein %>% mutate(Protein= as.factor(Chicken + 2*Beef + 3*Pork + 4*Shrimp + 5*Other + 6*Breakfast))
head(burrito_no_double_protein)

proteins = c('Chicken', 'Beef', 'Pork', 'Shrimp', 'Other', 'Breakfast')
proteins_X = as.matrix(burrito_no_double_protein[proteins])
head(proteins_X)

protein_cost_y = burrito_no_double_protein$Cost

```

```{r protein_model}
set.seed(RANDOM_SEED)

mlr_protein_post_dist = truncated_gibbs(proteins_X, protein_cost_y, mu=rep(1, 7), 
                                        tau_2, a, b, lb=rep(0,7), ub=rep(Inf,7))
mlr_protein_post_dist = mlr_protein_post_dist[5001:1000, ]
head(mlr_protein_post_dist)
```

## Model Diagnostics

```{r mlr_trace_plots}
plot_traces(mlr_protein_post_dist[,1:4], 'Parameter Traces (After Burn In)')
plot_traces(mlr_protein_post_dist[,5:8], 'Parameter Traces (After Burn In)')
```

```{r mlr_acf_plots}
acf_plots(mlr_protein_post_dist[,1:4])
acf_plots(mlr_protein_post_dist[,5:8])
```


```{r mlr_summary}
summarize_dist(mlr_protein_post_dist, colnames(mlr_protein_post_dist), round_places=2)
```

# Mixed Model

```{r mixed model}
set.seed(RANDOM_SEED)

a1 <- 0.1975
a2 <- 0.44
b1 <- 0.5
b2 <- 0.5

mixed_post = mixed_effects_gibbs(proteins_X, protein_cost_y, 
                                 group=burrito_no_double_protein$Location, 
                                 mu=rep(1, p), tau_2, a1, b1, a2, b2)

mixed_post <- mixed_post[5001:10000,]
head(mixed_post)
```

## Model Diagnostics

```{r mixed_trace_plots}
plot_traces(mixed_post[,1:4], 'Parameter Traces (After Burn In)')
plot_traces(mixed_post[,5:9], 'Parameter Traces (After Burn In)')
```

```{r mixed_acf_plots}
acf_plots(mixed_post[,1:4])
acf_plots(mixed_post[,5:9])
```

The ACF plots for $\sigma$ and $\kappa$ both have some auto correlation.  This is not desirable, but was expected after seeing some autocorrelation in these parameters on Homework 8.  It is ignored in this project.

```{r mixed_summary}
summarize_dist(mixed_post, colnames(mixed_post), round_places = 2)
```


# Model Comparisons

```{r Deviance Information Criterion}
mlr_dic <- dic(x=ingredient_X, 
               beta=mlr_post_dist[,-ncol(mlr_post_dist)], 
               sig2=mlr_post_dist[,ncol(mlr_post_dist)], 
               y=cost_y)

truncated_dic <- dic(x=ingredient_X, 
                     beta=truncated_post_dist[,-ncol(truncated_post_dist)], 
                     sig2=truncated_post_dist[,ncol(truncated_post_dist)], 
                     y=cost_y)

reduced_dic <- dic(x=proteins_X, 
                   beta=mlr_protein_post_dist[,-ncol(mlr_protein_post_dist)],
                   sig2=mlr_protein_post_dist[,ncol(mlr_protein_post_dist)],
                   y=protein_cost_y)

mixed_dic <- dic(x=proteins_X, 
                 beta=mixed_post[,1:7],
                 sig2=mixed_post[,8],
                 y=protein_cost_y)

mlr_dic
truncated_dic
reduced_dic
mixed_dic

```

# Model Interpretations

The reduced model has the smallest DIC of 77 compared to -152 in the MLR, -129 in the truncated MLR, and 172 in the mixed model.  According to DIC, this means that the reduced model best explains the variance in our data.


